---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# BOSS

<!-- badges: start -->
<!-- badges: end -->

The goal of BOSS is to perform scalable approximate Bayesian inference for complex hierarchical models, with a (log) marginal likelihood that is intractable and/or expensive to compute.

## Installation

You can install the development version of BOSS from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("davidolohowski/BOSS")
```

## Example

This is a basic example which shows you how to solve a simple problem in 1-D.

Assume the true (log) marginal likelihood is given by the following sinusoidal function:

```{r example}
library(BOSS)
f1d <- function(x) sin(0.5*pi*x)
```

We can use the `boss` function to approximate the (log) marginal likelihood of this function, specifying the objective function, the dimension of the problem, and the bounds for the BO optimization. 
The `alpha` parameter controls the level of the essential support, while `h` controls the target fill-in distance of the design points.

```{r}
# A automatic implementation
br2 <- boss(f1d, D = 1, 
            modal_opts = list(lower = 0, upper = 2),
            alpha = 0.05, h = 0.1, verbose = 1)
plot(br2)
```


For more advanced applications, the user can also manually perform each step of the BOSS algorithm, as shown below:
```{r}
# A manual implementation
br <- BOSS:::BOSS_modal(func = f1d, D = 1,
                        lower = 0, upper = 2,
                        update_step = 5,
                        quad = FALSE,
                        max_iter = 20,
                        verbose = 1)
br <- update_hessian(br)
br <- compute_essential_support(br, alpha=0.05)
br <- construct_essential_designs(br)
br <- compute_fill_in(br)
br <- fill_in(br, h = 0.1, verbose = 1, max_add = 100)
br_update <- update_boss(br)
plot(br_update)
```

## 2-D Example

Test the BOSS algorithm on a 2-D Gaussian density. The objective function here is the log-density of a randomly generated bivariate Gaussian random variable.

```{r exmple2d}
rho <- runif(1, min = -0.9, 0.9)
f2d <- function(x) mvtnorm::dmvnorm(x, rep(0,2), sigma = matrix(c(1,rho, rho, 1), 2, 2), log = T)

# A automatic implementation
br3 <- boss(f2d, D = 2, 
            modal_opts = list(lower = -c(3,3), upper = c(3,3)),
            alpha = 0.05, h = 0.1, verbose = 1)
plot(br3)
```

Similarly, users can also have more modular control over the procedure:

```{r}
br3 <- BOSS:::BOSS_modal(func = f2d, D = 2,
                        lower = -c(3,3), upper = c(3,3),
                        initial_design = 10,
                        quad = TRUE,
                        update_step = 10,
                        max_iter = 100,
                        verbose = 1)

br3 <- update_hessian(br3)
br3 <- compute_essential_support(br3, alpha = 0.05)
br3 <- construct_essential_designs(br3)
br3 <- compute_fill_in(br3, n_samples = 100000)
br3 <- fill_in(br3, h = 0.1, max_add = 100)
plot(br3)

br3_update <- update_boss(br3)
```



## More efficient surrogate

By default, the current functions `update_boss` and `boss_modal` construct the `surrogate` function using the function `predict_gp`, which implies that each evaluation of the surrogate function requires a full matrix inversion of the covariance matrix. 
This becomes inefficient if we want to evaluate the surrogate function at many points.

To improve the efficiency, we can use the `update_boss_faster` function, which uses pre-computed factorization of the covariance to define the surrogate function.
The `surrogate` function updated by `update_boss_faster` is much more efficient than the original `surrogate` function, especially when the number of design points is large.

```{r}
br3_update_faster <- update_boss_faster(br3)
```

```{r}
br3_update$surrogate(c(1,1))
br3_update_faster$surrogate(c(1,1))
```


```{r}
compare_runtime <- microbenchmark::microbenchmark(
  br3_update$surrogate(c(1,1)),
  br3_update_faster$surrogate(c(1,1)),
  unit = 'ms',
  times = 1000
)

compare_runtime
```


## More efficient BO

Each evaluation of the AF also involves computing all the GP-related quantities, which can be made more efficient in the same way as above.

```{r}
set.seed(123)
br3 <- BOSS:::BOSS_modal(func = f2d, D = 2,
                        lower = -c(3,3), upper = c(3,3),
                        initial_design = 10,
                        quad = TRUE,
                        update_step = 10,
                        max_iter = 100,
                        verbose = 1)

set.seed(123)
br3_v2 <- BOSS:::BOSS_modal_v2(func = f2d, D = 2,
                        lower = -c(3,3), upper = c(3,3),
                        initial_design = 10,
                        quad = TRUE,
                        update_step = 10,
                        max_iter = 100,
                        verbose = 1)
```


The two implementation are pragmatically the same in terms of the design points queried.

```{r}
mean(abs(br3$design_points$x - br3_v2$design_points$x))
```

```{r}
mean(abs(br3$design_points$y - br3_v2$design_points$y))
```

Similarly for the 1-D example:


```{r}
set.seed(123)
br2 <- BOSS:::BOSS_modal(func = f1d, D = 1,
                        lower = 0, upper = 2,
                        update_step = 5,
                        quad = FALSE,
                        max_iter = 20,
                        verbose = 1)

set.seed(123)
br2_v2 <- BOSS:::BOSS_modal_v2(func = f1d, D = 1,
                        lower = 0, upper = 2,
                        update_step = 5,
                        quad = FALSE,
                        max_iter = 20,
                        verbose = 1)
```

Check the two versions of the 1-D BOSS algorithm produce the same design points:

```{r}
mean(abs(br2$design_points$x - br2_v2$design_points$x))
```

```{r}
mean(abs(br2$design_points$y - br2_v2$design_points$y))
```


