---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# BOSS

<!-- badges: start -->
<!-- badges: end -->

The goal of BOSS is to perform scalable approximate Bayesian inference for complex hierarchical models, with a (log) marginal likelihood that is intractable and/or expensive to compute.

## Installation

You can install the development version of BOSS from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("davidolohowski/BOSS")
```

## Example

This is a basic example which shows you how to solve a simple problem in 1-D.

Assume the true (log) marginal likelihood is given by the following sinusoidal function:

```{r example}
library(BOSS)
f1d <- function(x) sin(0.5*pi*x)
```

We can use the `boss` function to approximate the (log) marginal likelihood of this function, specifying the objective function, the dimension of the problem, and the bounds for the BO optimization. 
The `alpha` parameter controls the level of the essential support, while `h` controls the target fill-in distance of the design points.

```{r}
# A automatic implementation
br2 <- boss(f1d, D = 1, 
            modal_opts = list(lower = 0, upper = 2),
            alpha = 0.05, h = 0.1, verbose = 1)
plot(br2)
```


For more advanced applications, the user can also manually perform each step of the BOSS algorithm, as shown below:
```{r}
# A manual implementation
br <- BOSS:::BOSS_modal(func = f1d, D = 1,
                        lower = 0, upper = 2,
                        update_step = 5,
                        max_iter = 20,
                        verbose = 1)
br <- update_hessian(br)
br <- compute_essential_support(br, alpha=0.05)
br <- construct_essential_designs(br)
br <- compute_fill_in(br)
br <- fill_in(br, h = 0.1, verbose = 1, max_add = 100)
br_update <- update_boss(br)
plot(br_update)
```



